### 2017-04-06 paper.tex
Add Serial Payload Bus

Before a protocol was chosen, two different preliminary designs approaches were
prototyped and

\subsubsection

>>>>

Before a protocol was chosen, two different preliminary designs were devised
and experimented with outside of the full flight platform to ensure they were
viable, and which of the two was worth developing further.

\subsubsection{Serial Bus Interface}

The first experimental bus interface was designed around a Serial interface.
Serial interfaces (based on \acrshort{uart}) are only meant to connect two
devices, and require two data wires (one for each transmission direction). To
avoid requiring a pair of input/output pins on the \acrlong{obc} for each
connected payload, and simplify the electronic layout, all payloads were wired
in parallels on the same reception (Rx) and transmission (Rx) data lines. When
a payload had any data ready to be sent, it would transmit it using
\acrshort{uart} to the \acrshort{obc}, which would then store it in memory.

This configuration proved unusable. Because \acrshort{uart} is not designed
to handle multiple devices on the same connection endpoint, data would be
corrupted whenever two or more payloads were attempting to send data to the
\acrshort{obc}.

The chosen solution for this problem was to make four bus lines available to
each payload. In addition to the common Rx/Tx lines, each interface is composed
of a Communication Request Line (CR) and a Communication Enable Line (CE). With
the new design, the communication sequence had to be changed to the following:

\begin{enumerate}
\item When a payload is ready to send data, it pulls its CR line high.
\item The \acrshort{obc} lists payloads that have raised CR. For each:
\begin{enumerate}
    \item The \acrshort{obc} pulls the payload's CE line high.
    \item The payload sends its data to the \acrshort{obc}.
    \item The \acrshort{obc} pulls the payload's CE line low.
    \item The payload pulls its CR line low.
    \item The \acrshort{obc} processes the received data.
\end{enumerate}
\end{enumerate}

To ensure that non-cooperative payload could not attempt to send data when the
\acrshort{obc} had not addressed them through pulling the CE line high, a logic
AND gate (two NAND gates chained together as shown in figure \ref{fig:uart-nor})
was placed between the payload's Tx line and its CE line. With this design,
any signal on the Tx line could not get past the gate unless the payload was
being addressed, thus preventing a payload from corrupting another's signal.

\begin{figure}[H]
\includegraphics[width=0.6\textwidth]{serial-nand.png}
\centering
\caption{Payload Tx line protection}
\label{fig:uart-nor}
\end{figure}

This solution was tested using a Raspberry Pi 3 for the \acrlong{obc}, and an
Arduino UNO board for the payload. An unforeseen electrical engineering issue
required slight redesign of the system: on both the Arduino or the Raspberry
Pi, a small voltage, even inferior to the logic high (\SI{3.3}{V} on Raspberry
Pi, \SI{5}{V} on the Arduino UNO) was enough for the line to be read as a
logical high. This caused certain bursts of noise in the lines to be detected
by the \acrshort{obc} as a high CR line, even though the payload had not
raised it. This was fixed by using inverted logic: a payload would signal 
available data by pulling the \(\overline{CE}\) line low, and the \acrshort{obc}
would pull \(\overline{CR}\) low when addressing a payload.

\begin{figure}[H]
\includegraphics[width=0.6\textwidth]{serial-wires.jpg}
\centering
\caption{Payload Serial Bus test setup}
\label{fig:uart-nor}
\end{figure}

The system worked with the experimental setup, and the \acrlong{obc} and the
OBC were able to exchange data. However, the design was deemed flawed because
of the hardware requirements: while the Rx and Tx lines are shared between all
payloads, each still requires two I/O pins on the \acrshort{obc} for the 
\(\overline{CE}\) and \(\overline{CR}\) lines. With most micro-controllers
exposing between 5 and 15 I/O pins, this would have strongly limited how many
payloads could have been connected at the same time.

\subsubsection{\acrshort{i2c} Bus Interface}
